\chapter{The Frequency Ranges Within the Human Range of Hearing}\label{appendix:frequency-table}
\begin{longtable}[h!]{| l | l|}
	\hline
		General Frequency Range & Description of Range \\ 
		\hline
		\multirow{8}{*}{<20Hz - 60Hz} & The lowest threshold of human hearing. \\ & This includes many frequencies that \\ & are felt and not heard, and \\ & provides the \say{rumble} feeling in \\ & music. This range gives \\ & much of music its power, \\ & and is typically known as \\ & \say{sub-bass.} \\
		\hline
		\multirow{12}{*}{60Hz - 250Hz} & This range determines the amount \\ &  of \say{warmth} and how full the sound \\ & is perceived to be. The \\ & notes fundamental to rhythm \\ & lives in this range, and \\ & and too much sound in \\ & this frequency range will result \\ & in the overall sound being \\ & \say{boomy,} or muddy-sounding \\ & and messy. It is \\ & otherwise known as the \say{bass} \\ & frequency. \\
		\hline
		\multirow{8}{*}{250Hz - 500Hz} & The lower harmonics of many \\ & instruments are in this range.  \\ & It is generally known as \\ & the \say{lower midrange} of frequencies, \\ & and can introduce listening \\ & fatigue and a telephone-quality \\ & to the sound if this range is \\ & emphasized too much. \\ 
		\hline
		\multirow{14}{*}{500Hz - 2 kHz} & This range is considered the \\ & middle of the midrange. It \\ & gives many instruments a sense of \\ & prominence in a mix, and \\ & determines how audible one instrument \\ & or vocalist is in comparison \\ & to another. If this range is \\ & emphasized, audio output may \\ & sound tinny and small, which \\ & could lead to listening \\ & and ear fatigue, as the \\ & human ear is sensitive to \\ & the human voice, and the \\ & frequencies it covers. \\ 
		\hline
		\multirow{8}{*}{2 kHz - 4 kHz} & The \say{upper midrange} is \\ & responsible for much of the \\ & attack sounds on percussive \\ & and rhythmic instruments. This range \\ & may add presence to the \\ & mix if boosted, but if \\ & it is emphasized too \\ & much, it may mask some \\ & speech recognition sounds. \\ & Listening fatigue may also \\ & set in if this range \\ & is emphasized too much, \\ & as the slightest boost in \\ & this range will result in a \\ & noticeable change in the \\ & sound's timbre. \\ 
		\hline
		\multirow{13}{*}{4 kHz - 6 kHz} & This range is known as the \\ & \say{presence} range. It defines a sound's \\ & clarity and the definition of voices \\ & and instruments that are present. \\ & If this range is boosted, \\ & instruments and voices may \\ & sound physically closer to the \\ & listener, and vice versa, with \\ & reducing this range, causing \\ & instruments and voices to sound \\ & further away. However, if this range is \\ & emphasized too much, a harsh, \\ & irritating sound may occur. \\
		\hline
		\multirow{12}{*}{6 kHz - 20 kHz} & This range controls the \say{brilliance} \\ & and clarity of sounds within the \\ & mix. Instead of pitches, this range is \\ & composed entirely of harmonics, and brings \\ & \say{sparkle} to the sound. This range also may \\ & easily cause ear fatigue, as an \\ & over-emphasis can increase the \\ & hiss heard, and produce sibilance, or \\ & an unpleasant tonal harshness, which \\ & can happen with consonant syllables \\ & (most noticeably: S, T, and Z), \\ & especially on vocals. \\
		\hline
\caption{The general frequency range, within the range of human hearing} \cite{Suits_1998}\cite{Zjalic_2021}
\label{tbl:frequency-table-of-human-hearing-general}
\end{longtable}